Beliefs about the development of mental life

### 2021-08-08 supplement/supplement-main.Rmd
edits for submission, add figs to readme

---
title: "Beliefs about the development of mental life"
subtitle: "Supplemental Materials"
author:
  - Kara Weisman^[Department of Psychology, University of California, Riverside; kgweisman@gmail.com]
  - Lucy S. King^[Department of Psychiatry and Behavioral Sciences, Tulane University School of Medicine; Department of Psychology, The University of Texas at Austin]
  - Kathryn L. Humprheys^[Department of Psychology, Vanderbilt University]
date: "August 7, 2021"
output:
  # word_document:
  #   reference_docx: word-styles-ref.docx
   bookdown::pdf_document2:
    latex_engine: xelatex
    fig_caption: yes
    toc: true
    toc_depth: 3
bibliography: "references.bib"
cls: article.cls
figsintext: yes
header-includes: 
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{float}
tables: yes
---

@@

\newpage
In a series of large-scale studies we assessed how US adults conceptualize the development of the human mind over the first five years of life. We identified four distinct suites of mental capacities that are perceived by US adults to develop over human infancy and early childhood, _bodily sensations_, _negative affect_, _social connection_, and _cognition and control_ (Study 1); charted how these different aspects of mental life are perceived to change over the first five years of a child’s life (Studies 2-3); and explored the lay theories that underlie the observed differences in the perceived developmental trajectories of these four aspects of mental life (Study 3).

In these Supplemental Materials, we provide extended information about the results presented in the main paper, present additional results, and provide an item-by-item comparison of the capacity items used here to items used in previous work on mind perception [@gray.etal2007; @weisman.etal2017a; @weisman.etal2017b; @malle2019].

@@

>>>>

---
title: "Beliefs about the development of mental life"
subtitle: "Supplemental Materials"
author:
  - Kara Weisman^[Department of Psychology, University of California, Riverside; kgweisman@gmail.com]
  - Lucy S. King^[Department of Psychiatry and Behavioral Sciences, Tulane University School of Medicine; Department of Psychology, The University of Texas at Austin]
  - Kathryn L. Humprheys^[Department of Psychology, Vanderbilt University]
date: "August 7, 2021"
output:
  # word_document:
  #   reference_docx: word-styles-ref.docx
  bookdown::pdf_document2:
    latex_engine: xelatex
    fig_caption: yes
    toc: true
    toc_depth: 2
    number_sections: false
bibliography: "references.bib"
cls: article.cls
figsintext: yes
header-includes: 
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{float}
tables: yes
---

@@

\newpage
In a series of large-scale studies we assessed how US adults conceptualize the development of the human mind over the first five years of life. We identified four distinct suites of mental capacities that are perceived by US adults to develop over human infancy and early childhood, _bodily sensations_, _negative affect_, _social connection_, and _cognition and control_ (Study 1); charted how these different aspects of mental life are perceived to change over the first five years of a child's life (Studies 2-3); and explored the lay theories that underlie the observed differences in the perceived developmental trajectories of these four aspects of mental life (Study 3).

In these Supplemental Materials, we provide extended information about the methods and results presented in the main paper, present additional results, and provide an item-by-item comparison of the capacity items used here to items used in previous work on mind perception [@gray.etal2007; @weisman.etal2017a; @weisman.etal2017b; @malle2019].

@@

# Extended methods

```{r}
# NOTE: this is a bit of a circular process -- results files are generated in this .Rmd file, but they also need to be loaded in order to write the Extended Methods section. For this reason, results RDS files are included in the github repo.
s1_results <- readRDS("../outputs/s1_results.rds")
s2_results <- readRDS("../outputs/s2_results.rds")
s3_results <- readRDS("../outputs/s3_results.rds")

# ALSO NOTE: if you encounter problems knitting, try clearing the knit cache.
```

Methods, inclusion/exclusion criteria, and analyses marked as "preregistered" were preregistered on the Open Science Framework (OSF) website: (Study 1: https://osf.io/e6ajh/; Study 2: https://osf.io/j72dg; Study 3: https://osf.io/xh8ce/). All studies were approved by the Stanford University Internal Review Board.

### Participants

`r s1_results$demo %>% distinct() %>% nrow()` US adults participated via Amazon Mechanical Turk (MTurk) in July–August 2018. To be eligible for the study, participants were required to have gained approval for at least 95% of their previous work on MTurk, with at least 50 HITs completed; have MTurk accounts based in the US; indicate that they were between the ages of 18-45 years; and report that their English abilities were either "advanced" or "superior" (not "novice," "intermediate," or decline to provide information about English ability). Participants were paid $2.00 for approximately 10-15 minutes of their time.

According to self-report, participants ranged in age from `r min(s1_results$demo$Age)`-`r max(s1_results$demo$Age)` years ($M$ = `r round_x(mean(s1_results$demo$Age, na.rm = T), 2)` years, $sd$ = `r round_x(sd(s1_results$demo$Age, na.rm = T), 2)` years) and included more men (`r round((s1_results$demo %>% count(GenderSex) %>% mutate(prop = n/sum(n)) %>% filter(GenderSex == "Male") %>% pull(prop))*100, 0)`%) than women (`r round((s1_results$demo %>% count(GenderSex) %>% mutate(prop = n/sum(n)) %>% filter(GenderSex == "Female") %>% pull(prop))*100, 0)`%; <1% of participants identified as some other gender or declined to disclose). Participants predominantly identified as White (`r round((s1_results$demo %>% count(RaceEthnicity) %>% mutate(prop = n/sum(n)) %>% filter(RaceEthnicity == "White") %>% pull(prop))*100, 0)`%; <`r round(100*(s1_results$demo %>% count(RaceEthnicity) %>% mutate(prop = n/sum(n)) %>% filter(RaceEthnicity != "White") %>% filter(prop == max(prop)) %>% pull(prop)) + 1, 0)`% identified as any other race/ethnicity, identified as more than one race/ethnicity, or declined to disclose). `r round(100*(s1_results$demo %>% count(Education) %>% arrange(desc(Education)) %>% mutate(cumprop = cumsum(n)/sum(n)) %>% filter(grepl("Bachelor", Education)) %>% pull(cumprop)))`% of participants had obtained at least a Bachelor's degree. `r round(100*(s1_results$demo %>% count(Parent) %>% mutate(prop = n/sum(n)) %>% filter(Parent == "Yes") %>% pull(prop)))`% of participants indicated that they were parents.

An additional 77 participants completed the study but were excluded for failing one or more (out of three) embedded attention checks (e.g., failing to choose "34" in response to the question "Please select 34") or for indicating via a multiple choice question at the end of the survey that they had not paid attention, had not avoided distractions, or had not taken the study tasks seriously.

### Materials and procedure

After completing a screening questionnaire (in which participants provided information about their gender/sex, age, and English abilities) and providing informed consent, participants were provided with an overview of the study, as follows: "In this study, we will ask you about the mental capacities of children of different ages, from newborns to 5-year-olds. We will show you example photos of children at three of the following ages to remind you of what children look like at that age, but please think about children of that age in general when answering these questions. There are 60 brief questions for each of the three ages (for example, 'To what extent is a newborn capable of feeling happy?'), followed by a few demographic questions about yourself. We are interested in your opinions about these questions. Please do not look up extra information about these questions—we just want to know what you think." Participants were shown an array of pictures of children at several target ages between 0-5 years; see "Target ages and accompanying photographs," below.

Participants then proceeded through three trials in which they assessed the mental lives of children at three target ages: birth, 9 months, and 5 years. On each trial, they were shown two representative photographs of children at the target age (with the photograph labeled "newborns," "9-month-olds," or "5-year-olds"), and asked to answer the following question for 60 capacities: "To what extent is a [newborn/9-month-old/5-year-old] capable of [this capacity]?" Participants responded on a sliding scale from 0 (labeled as "not at all capable") to 100 (labeled as "completely capable").

On the final page of the survey, participants were asked to provide additional demographic information.

#### Capacities

The 60 capacities included in Study 1 were drawn from several areas of previous research. Our goals in selecting items were to: (1) balance the representation of the conceptual organization underlying mental capacity attributions in general, as identified in previous work [@gray.etal2007; @weisman.etal2017a; @weisman.etal2017b; @weisman.etal2018; @malle2019]; (2) add items relevant to early development, as identified by existing measures of temperament and behavior in early life [@putnam.etal2006; @rothbart1978]; and (3) assess the broadest range of mental capacities within financial and participant burden constraints, as identified by extensive discussions among the authors drawing on our own research and clinical experience with infants and young children.

This process yielded a list of 60 capacities. Roughly half of this list (29 of 60 items) were drawn from (or closely based on) the items included in Weisman et al.'s recent studies [@weisman.etal2017a]: "being aware of things", "feeling calm", "feeling embarrassed", "feeling guilty", "feeling happy", "feeling love", "feeling pain", "feeling pleasure", "feeling pride", "feeling sad", "feeling safe", "feeling scared", "feeling tired", "getting angry", "getting hungry", "getting hurt feelings", "having goals", "having self-control", "having thoughts", "having wants and desires", "hearing sounds", "making choices", "planning", "reasoning about things", "recognizing others' emotions", "recognizing somebody else", "remembering things", "seeing", and "telling right from wrong". By extension, the list included close variants on 14 of the 20 capacities evaluated in Gray et al.'s original study [@gray.etal2007].

The other half (31 of 60 items) were novel items, not used in previous studies of mind perception, but common in clinical assessments of infant temperament and behavior: "being afraid of somebody", "being angry at somebody", "being comforted by physical touch", "calming themselves down", "controlling their emotions", "detecting danger", "feeling annoyed", "feeling bored", "feeling confused", "feeling distressed", "feeling excited", "feeling frustrated", "feeling helpless", "feeling hopeless", "feeling lonely", "feeling neglected", "feeling overwhelmed", "feeling physically uncomfortable", "feeling textures (for example, 'smooth', 'rough')", "feeling thirsty", "feeling too hot or too cold", "feeling worried", "finding something funny", "focusing on a goal", "getting pleasure from music", "imagining things", "learning from other people", "listening to somebody", "loving somebody", "thinking before they act", and "understanding what somebody else is thinking".

(See \@ref(label:"Selection of capacities: Comparison to previous studies")  for an item-by-item comparison of the capacities used here to those used in previous studies of mind perception [@gray.etal2007; @weisman.etal2017a; @weisman.etal2017b; @malle2019].)

Capacities were presented in a random order each time they were displayed (i.e., order varied across both target ages and participants).

#### Target ages and accompanying photographs

In planning this series of studies, we selected 13 target ages based on the developmental milestone ages identified by the CDC [@CDC2021] with the addition of the American Academy of Pediatrics' recommendations for pediatrician visits before 2 months [@AAP2018]: birth, 4 days, 1 month, 2 months, 4 months, 6 months, 9 months, 12 months, 18 months, 2 years, 3 years, 4 years, and 5 years. We limited Study 1 to just the youngest, median, and oldest target ages—presented to participants as "newborns," "9-month-olds" and "5-year-olds"—in order to maximize the number of capacities included in the study without over-burdening participants.

In selecting photos for each target age, we aimed to have photos of the same (or very similar-looking) children at all target ages so as to emphasize the question of the development of mental capacities over the course of an individual child's early life. We chose to illustrate each target age with two photos in order to emphasize that we were interested in children's mental capacities in general, rather than the experiences of a particular child in a particular moment. We also presented participants with instructions that underscored this point (see above). We included one boy and one girl at each target age in order to generalize across perceived sex/gender differences in children's early mental life. We used photographs of two (apparently) White, non-Hispanic children, for three reasons: (1) In order to avoid introducing spurious differences across target ages by confounding a target's age and race/ethnicity; (2) In order to match the anticipated race/ethnicity of the majority of participants on MTurk; and (3) Because of the limited set of high-resolution photographs of individual children of color at different points of development currently available to us. Conducting similar studies of adults' perceptions of the development of mental life with more ethnically and racially diverse participants, and examining the effects of sex, gender, race, and ethnicity on adults' perceptions of children's mental life, will be critical next steps in this line of research.

Target ages were presented in chronological order, with participants first assessing all 60 capacities for newborns, then 9-month-old infants, and finally 5-year-old children.

### Analysis plan

To conduct the EFAs reported in Results, we proceeded as if each participant's ratings of the 60 mental capacities for each of the three target ages were a distinct source of data; in other words, as if we had 903 participants assessing one target age each, instead of 301 participants assessing three target ages each. (See [@weisman.etal2017a; @weisman.etal2021], for similar treatments of within-subjects data in the context of EFA.) We used Pearson correlations to obtain minimal residual solutions and applied an oblique transformation (oblimin) to the resulting factor structures, using the "psych" package for R [@revelle2021].

In order to determine how many factors to retain, we examined the results of three factor retention protocols: (1) Parallel analysis, which compares the observed correlation structure to the correlation structure arising from random datasets of the same size; (2) Minimizing the Bayesian Information Criterion (BIC), which is one method of optimizing both goodness of fit and parsimony; and (3) A set of factor retention criteria that have been used in Weisman et al.'s previous work [@weisman.etal2017a], in which they retained factors with eigenvalues greater than 1.0, which individually accounted for greater than 5% of the shared variance before transformation or rotation, and which were the "dominant" factor (the factor with the strongest absolute factor loading) for at least one mental capacity after transformation (a factor had to meet all three of these criteria in order to be retained). Our interpretation of how best to characterize our dataset (i.e., how many factors we observe) was guided by the degree of consensus among these three protocols and the interpretability of the retained factors under each protocol.

Parallel analysis suggested a `r s1_results$efa$factors`-factor solution (accounting for `r round(100*(s1_results$efa$Vaccounted["Cumulative Var",ncol(s1_results$efa$Vaccounted)]))`% of the total variance); minimizing BIC suggested a `r s1_results$efa_bic$factors`-factor solution (accounting for `r round(100*(s1_results$efa_bic$Vaccounted["Cumulative Var",ncol(s1_results$efa_bic$Vaccounted)]))`% of the total variance); and Weisman et al.'s (2017) factor retention criteria suggested a `r s1_results$efa_wdm$factors`-factor solution (accounting for `r round(100*(s1_results$efa_wdm$Vaccounted["Cumulative Var",ncol(s1_results$efa_wdm$Vaccounted)]))`% of the total variance). We present the `r s1_results$efa$factors`-factor solution in the main text because it was of intermediate size and because we judged the factors it revealed to be intelligible and of theoretical interest; we present all three solutions in these Supplemental Materials.

### Participants

`r s2_results$demo %>% distinct() %>% nrow()` US adults participated via MTurk in August 2018. Eligibility requirements were identical to Study 1. Participants were paid $2.90 for approximately 15-20 minutes of their time. In response to the overrepresentation of men in Study 1, in Study 2 we recruited men and women through separate "Human Intelligence Tasks" (HITs).

According to self-report, participants ranged in age from `r min(s2_results$demo$Age)`-`r max(s2_results$demo$Age)` years ($M$ = `r round_x(mean(s2_results$demo$Age, na.rm = T), 2)` years, $sd$ = `r round_x(sd(s2_results$demo$Age, na.rm = T), 2)` years) and (per our recruitment strategy) included roughly equal numbers of men (`r round((s2_results$demo %>% count(GenderSex) %>% mutate(prop = n/sum(n)) %>% filter(GenderSex == "Male") %>% pull(prop))*100, 0)`%) and women (`r round((s2_results$demo %>% count(GenderSex) %>% mutate(prop = n/sum(n)) %>% filter(GenderSex == "Female") %>% pull(prop))*100, 0)`%). Participants predominantly identified as White (`r round((s2_results$demo %>% count(RaceEthnicity) %>% mutate(prop = n/sum(n)) %>% filter(RaceEthnicity == "White") %>% pull(prop))*100, 0)`%; <`r round(100*(s2_results$demo %>% count(RaceEthnicity) %>% mutate(prop = n/sum(n)) %>% filter(RaceEthnicity != "White") %>% filter(prop == max(prop)) %>% pull(prop)) + 1, 0)`% identified as any other race/ethnicity, identified as more than one race/ethnicity, or declined to disclose). `r round(100*(s2_results$demo %>% count(Education) %>% arrange(desc(Education)) %>% mutate(cumprop = cumsum(n)/sum(n)) %>% filter(grepl("Bachelor", Education)) %>% pull(cumprop)))`% of participants had obtained at least a Bachelor's degree. `r round(100*(s2_results$demo %>% count(Parent) %>% mutate(prop = n/sum(n)) %>% filter(Parent == "Yes") %>% pull(prop)))`% of participants indicated that they were parents.

An additional 90 participants completed the study but were excluded for failing one or more (out of four) attention checks or indicating that they had not paid attention (see Study 1).

### Materials and procedure

After completing a screening questionnaire (in which participants provided information about their gender/sex, age, and English abilities) and providing informed consent, participants were provided with an overview of the study that was nearly identical to that used in Study 1.

Participants then proceeded through 13 trials in which they assessed the mental lives of children at 13 target ages: birth, 4 days, 1 month, 2 months, 4 months, 6 months, 9 months, 12 months, 18 months, 2 years, 3 years, 4 years, and 5 years. On each trial, they were shown two representative photographs of children at the target age, and asked to answer the following question for 20 capacities: "To what extent is a [newborn/4-day-old/etc.] capable of [this capacity]?" As in Study 1, participants responded on a sliding scale from 0 (labeled as "not at all capable") to 100 (labeled as "completely capable").

On the final page of the survey, participants were asked to provide additional demographic information.

#### Capacities

In order to ask participants to assess a more fine-grained array of target ages without undue participant burden, we limited our list of capacities to 20 of the 60 capacities used in Study 1. To identify this list, we drew on the results of Study 1, selecting 5 items for each of the 4 factors identified there. For each factor, we aimed to select items that loaded strongly on that factor, did not cross-load strongly on other factors, were sufficiently distinguishable from each other in meaning (i.e., not overly redundant), and captured our qualitative understanding of the latent construct that each factor corresponded to.

To represent _bodily sensations_, we chose the following items: "getting hungry", "feeling pain", "feeling tired", "feeling physically uncomfortable", and "hearing sounds". In the oblimin-transformed `r s1_results$efa$factors`-factor solution from Study 1 all of these items had loadings $\geq$ `r s1_results$efa$loadings[] %>% data.frame() %>% rownames_to_column("capacity") %>% filter(capacity %in% c("getting_hungry", "feeling_pain", "feeling_tired", "feeling_physically_uncomfortable", "hearing_sounds")) %>% pull("MR2") %>% min() %>% round_x(., 2)` on this factor, and absolute loadings $\leq$ `r s1_results$efa$loadings[] %>% data.frame() %>% rownames_to_column("capacity") %>% filter(capacity %in% c("getting_hungry", "feeling_pain", "feeling_tired", "feeling_physically_uncomfortable", "hearing_sounds")) %>% gather(factor, loading, -capacity) %>% filter(factor != "MR2") %>% pull(loading) %>% max((abs(.))) %>% round_x(., 2)` on the other three factors.

To represent _negative affect_, we chose the following items: "feeling distressed", "feeling lonely", "feeling frustrated", "feeling helpless", and "feeling overwhelmed". In Study 1 all of these items had loadings $\geq$ `r s1_results$efa$loadings[] %>% data.frame() %>% rownames_to_column("capacity") %>% filter(capacity %in% c("feeling_distressed", "feeling_lonely", "feeling_frustrated", "feeling_helpless", "feeling_overwhelmed")) %>% pull("MR4") %>% min() %>% round_x(., 2)` on this factor, and absolute loadings $\leq$ `r s1_results$efa$loadings[] %>% data.frame() %>% rownames_to_column("capacity") %>% filter(capacity %in% c("feeling_distressed", "feeling_lonely", "feeling_frustrated", "feeling_helpless", "feeling_overwhelmed")) %>% gather(factor, loading, -capacity) %>% filter(factor != "MR4") %>% pull(loading) %>% max((abs(.))) %>% round_x(., 2)` on the other three factors. As these numbers suggest, this was a difficult factor to represent, because there were no items that loaded very strongly on the factor (highest loading: `r s1_results$efa$loadings[] %>% data.frame() %>% rownames_to_column("capacity") %>% filter(capacity %in% c("feeling_distressed", "feeling_lonely", "feeling_frustrated", "feeling_helpless", "feeling_overwhelmed")) %>% pull("MR4") %>% max() %>% round_x(., 2)`) and because the highest-loading items tended to cross-load on other factors as well. Nonetheless, we decided to try to represent this latent construct because of its theoretical interest and potential clinical importance.

To represent _social connection_, we chose the following items: "feeling excited", "finding something funny", "loving somebody", "learning from other people", and "feeling happy". In the oblimin-transformed `r s1_results$efa$factors`-factor solution from Study 1 all of these items had loadings $\geq$ `r s1_results$efa$loadings[] %>% data.frame() %>% rownames_to_column("capacity") %>% filter(capacity %in% c("feeling_excited", "finding_something_funny", "loving_somebody", "learning_from_other_people", "feeling_happy")) %>% pull("MR3") %>% min() %>% round_x(., 2)` on this factor, and absolute loadings $\leq$ `r s1_results$efa$loadings[] %>% data.frame() %>% rownames_to_column("capacity") %>% filter(capacity %in% c("feeling_excited", "finding_something_funny", "loving_somebody", "learning_from_other_people", "feeling_happy")) %>% gather(factor, loading, -capacity) %>% filter(factor != "MR3") %>% pull(loading) %>% max(abs(.)) %>% round_x(., 2)` on the other three factors.

Finally, to represent _cognition and control_, we chose the following items: "planning", "having self-control", "reasoning about things", "controlling their emotions", and "telling right from wrong". In the oblimin-transformed `r s1_results$efa$factors`-factor solution from Study 1 all of these items had loadings $\geq$ `r s1_results$efa$loadings[] %>% data.frame() %>% rownames_to_column("capacity") %>% filter(capacity %in% c("planning", "having_self_control", "reasoning_about_things", "controlling_their_emotions", "telling_right_from_wrong")) %>% pull("MR1") %>% min() %>% round_x(., 2)` on this factor, and absolute loadings $\leq$ `r s1_results$efa$loadings[] %>% data.frame() %>% rownames_to_column("capacity") %>% filter(capacity %in% c("planning", "having_self_control", "reasoning_about_things", "controlling_their_emotions", "telling_right_from_wrong")) %>% gather(factor, loading, -capacity) %>% filter(factor != "MR1") %>% pull(loading) %>% max(abs(.)) %>% round_x(., 2)` on the other three factors.

Capacities were presented in a random order each time they were displayed (i.e., order varied across both target ages and participants).

#### Target ages and accompanying photographs

The target ages for this study (birth, 4 days, 1 month, 2 months, 4 months, 6 months, 9 months, 12 months, 18 months, 2 years, 3 years, 4 years, and 5 years) were based on the developmental milestone ages described in Study 1. The pairs of photos used to illustrate each target age were identical to those used in Study 1; see Study 1, Methods, for more information.

As in Study 1, target ages were presented in chronological order, with participants first assessing all 20 capacities for newborns, then 4-day-old infants, then 1-month-old infants, and so forth.

### Analysis plan

#### Preregistered confirmatory analysis of conceptual structure

Based on the results of Study 1, we predicted that the correlational structure of participants' responses would reveal four underlying constructs corresponding to the four factors we identified in Study 1: _bodily sensations_, _negative affect_, _social connection_, and _cognition and control_. To test this hypothesis, we performed an EFA, following the methods established in Study 1. We predicted that, using the same guidelines we applied for conducting and interpreting EFA in Study 1, we would find support for a `r s1_results$efa$factors`-factor solution, with factors that were similar to the factors revealed in that study (e.g., similar groups of capacities that loaded strongly on any given factor).

In line with our predictions, both parallel analysis and Weisman et al.'s factor retention criteria suggested a `r s2_results$efa$factors`-factor solution, which accounted for `r round(100*(s2_results$efa$Vaccounted["Cumulative Var",ncol(s2_results$efa$Vaccounted)]))`% of the total variance in participants' capacity attributions. In this solution, all items loaded most strongly on the factors that they were selected to represent. Minimizing BIC suggested an `r s2_results$efa_bic$factors`-factor solution (accounting for `r round(100*(s2_results$efa_bic$Vaccounted["Cumulative Var",ncol(s2_results$efa_bic$Vaccounted)]))`% of the total variance).

#### Preregistered confirmatory analysis of perceived developmental trajectories

We preregistered two hypotheses about the perceived developmental trajectories we anticipated to observe in Study 2. (1) We predicted that, on average, participants' perception of children's mental capacities would increase as a function of the target child's age, with participants generally attributing more and greater abilities to older children; and (2) We predicted that, in the domains of _cognition and control_ and _social connection_, participants would perceive relatively more dramatic developmental changes extending later into childhood (i.e., lower intercepts and steeper slopes), whereas in the domains of _bodily sensations_ and _negative affect_ participants would perceive relatively smaller changes across development (i.e., higher intercepts and shallower slopes).

To test these hypotheses, we conducted a multilevel beta regression via a generalized additive model, regressing item-level responses onto the domain (i.e., which of the four "factors" this capacity was selected to represented) and a spline-based smooth of the target age of the child fitted separately for each domain, including the maximal random effects structure that would allow the model to converge. Using the "mgcv" package for R [@wood2017], we specified this model as follows: _gamm(response ~ s(target age in years, by = domain) + domain, random = list(subject ID = ~1, capacity = ~1)_. Attempting to model random slopes for domain or for target age resulted in model convergence problems. Responses were rescaled to range from 0.005 to 0.995 in order to satisfy the constraints of the beta distribution. Domain was dummy-coded with _cognition and control_ as the baseline.

These tests deviated slightly from the preregistered analysis. We originally planned to use "factor scores" (which take into account participants' responses to all 20 items and the relations of those items to all 4 factors) rather than the summary scores described above as our dependent variable for these tests. We planned to derive these factor scores by projecting Study 2 data into the conceptual space defined by the factor solution for Study 1; however, this projection was impossible to implement, because Study 2 included only a subset of the variables used in Study 1. Furthermore, upon reflection, we decided that using any kind of factor scores (even factor scores derived only from Study 2 data) would be inappropriate for testing our hypotheses, which hinge on assessing the _absolute_ degree to which participants attributed different aspects of mental life to children of different ages; factor scores, which function much like z-scores, are instead designed to capture individual differences in responses (i.e., the degree to which a given participant attributed relatively greater or lesser capacities compared to the average of all participants).

In addition, we note that we originally planned to use a multilevel linear regression to analyze these scores. However, we now consider it misguided to model participants' responses using gaussian distribution, because—both in principle and in our empirical dataset—capacity ratings behaved asymptotically, plateauing near the ceiling of the scale (100, labeled "completely capable") at some point in development. The approach described here—using the beta distribution—is a far more accurate representation of these responses. Using generalized additive models via the "mgcv" package for R [@wood2017] allowed us to fine-tune the shape of these perceived developmental trajectories via spline-based smoothing (analogous to fitting polynomial terms in the context of linear regression, but without specifying in advance the order of the polynomials included in the model).

### Participants

`r s3_results$demo %>% distinct() %>% nrow()` US adults participated via MTurk in April 2019. Eligibility requirements were identical to Studies 1-2. Participants were paid $2.90 for approximately 15-20 minutes of their time. As in Study 2, we recruited men and women separately.

According to self-report, participants ranged in age from `r min(s3_results$demo$Age)`-`r max(s3_results$demo$Age)` years ($M$ = `r round_x(mean(s3_results$demo$Age, na.rm = T), 2)` years, $sd$ = `r round_x(sd(s3_results$demo$Age, na.rm = T), 2)` years) and (per our recruitment strategy) included roughly equal numbers of men (`r round((s3_results$demo %>% count(GenderSex) %>% mutate(prop = n/sum(n)) %>% filter(GenderSex == "Male") %>% pull(prop))*100, 0)`%) and women (`r round((s3_results$demo %>% count(GenderSex) %>% mutate(prop = n/sum(n)) %>% filter(GenderSex == "Female") %>% pull(prop))*100, 0)`%). Participants predominantly identified as White (`r round((s3_results$demo %>% count(RaceEthnicity) %>% mutate(prop = n/sum(n)) %>% filter(RaceEthnicity == "White") %>% pull(prop))*100, 0)`%; <`r round(100*(s3_results$demo %>% count(RaceEthnicity) %>% mutate(prop = n/sum(n)) %>% filter(RaceEthnicity != "White") %>% filter(prop == max(prop)) %>% pull(prop)) + 1, 0)`% identified as any other race/ethnicity, identified as more than one race/ethnicity, or declined to disclose). `r round(100*(s3_results$demo %>% count(Education) %>% arrange(desc(Education)) %>% mutate(cumprop = cumsum(n)/sum(n)) %>% filter(grepl("Bachelor", Education)) %>% pull(cumprop)))`% of participants had obtained at least a Bachelor's degree. `r round(100*(s3_results$demo %>% count(Parent) %>% mutate(prop = n/sum(n)) %>% filter(Parent == "Yes") %>% pull(prop)))`% of participants indicated that they were parents.

An additional 79 participants completed the study but were excluded for failing two or more (out of eight) embedded attention checks or for failing to provide a reasonable response in well-formed English sentences to an open-ended comprehension check included at the end of the survey.

### Materials and procedure

After completing a screening questionnaire (in which participants provided information about their gender/sex, age, and English abilities) and providing informed consent, participants were provided with an overview of the study, as follows: "In this study, we will ask you about the mental capacities of children of different ages. We will show you example photos of children at each of the following ages to remind you of what children look like at that age, but please think about children of that age in general when answering these questions. We will ask you about 8 different mental capacities: feeling happy, getting hungry, controlling emotions, feeling distressed, feeling pain, reasoning about things, learning from other people, and feeling helpless. For each mental capacity, we will ask you to assess the extent to which newborns (at birth) have this mental capacity, then 4-day-olds, 1-month-olds, 2-month-olds, 4-month-olds, and so on, ending with 5-year-olds. We'll also ask for your opinion about the different factors that might influence the development of these capacities, such as the child being ‘preprogrammed' to have certain abilities, or the child learning from the people or objects around him or her. At the end of the survey, we will ask you a few demographic questions about yourself. We are interested in your opinions about these questions. Please do not look up extra information about these questions—we just want to know what you think."

Participants then proceeded through eight trials, in which they assessed each of the eight capacities in turn; the order of capacities was randomized across participants. On each trial, participant assessed children at all 13 target ages used in Study 2 in a fixed order from youngest to oldest. Each age was presented along with a representative pair of photographs (also used in Studies 1 and 2), and participants were asked to answer the following question: "To what extent is a [newborn/4-day-old/1-month-old/etc.] capable of [this capacity]?" Participants responded on a sliding scale from 0 (labeled as "not at all capable") to 100 (labeled as "completely capable").

On each trial, at the bottom of the page participants completed a set of questions about the possible mechanisms that drive development for the capacity in question. First, they were asked, "In your opinion, how important is each of the following factors in the development of children's capacity for [capacity]?" They provided independent ratings on a Likert-type scale from 0 (labeled as "not at all important") to 6 (labeled as "extremely important"), for the following 10 developmental mechanisms: "the child is biologically 'preprogrammed' to have this ability"; "the child has experiences in the womb that given them this ability"; "the child's body grows and matures (for example, muscles get stronger, child gets taller)"; "the child's brain changes (for example, brain grows bigger, more or fewer connections between neurons)"; "the child's senses improve (for example, vision gets sharper, hearing improves)"; "the child observes the objects and the physical world around him or her"; "the child observes the people around him or her"; "the child interacts with the people around him or her"; "people explicitly teach the child how to do this"; "the child actively experiments with how to do this". These items were presented in a fixed order (roughly, from what we perceived to be the most innate or biological mechanisms of development, to what we perceived to be the most learned or social). They were then provided an opportunity to write in "any factors that play a role in the development of children's capacity for [capacity]". Finally, they were asked to choose among these 10 possible mechanisms (or to write in another mechanism) to answer the question, "If you had to choose just one, which of the following factors is the most important in the development of capacities for [capacity]?"
On the final page of the survey, participants were asked to provide additional demographic information.

#### Capacities

Because Study 3 included additional questions probing participants' theories of developmental mechanisms, we reduced the set of mental capacities from 20 in Study 2 to eight in the current study to maintain an appropriate level of participant burden. To select these eight capacities, we drew on the results of Study 2, selecting two items for each of the four factors identified by EFA. For each domain, we selected items that loaded strongly on the relevant factor, did not cross-load strongly on other factors, were sufficiently distinguishable from each other in meaning (i.e., not overly redundant), and captured our qualitative understanding of the domain that each factor corresponded to. We chose "controlling their emotions" and "reasoning about things" to represent the domain of _cognition and control_; "getting hungry" and "feeling pain" to represent the domain of _bodily sensations_; "learning from other people" and "feeling happy" to represent the domain of _social connection_; and "feeling distressed" and "feeling helpless" to represent the domain of _negative affect_. Capacities were presented in a random order for each participant.

#### Target ages and accompanying photographs

The 13 target ages for this study were identical to Study 2, with the exception that we added the parenthetical phrase "(at birth)" when referring to newborns. Each target age was presented with the same pair of photographs used in Study 2. Target ages were presented in chronological order on each trial.

### Analysis plan

#### Preregistered confirmatory analysis of perceived developmental trajectories

We preregistered two hypotheses about the perceived developmental trajectories we anticipated to observe in Study 3; these were identical to our hypotheses for Study 2 (which were upheld in that study): (1) We predicted that, on average, participants' perception of children's mental capacities would increase as a function of the target child's age, with participants generally attributing more and greater abilities to older children; and (2) We predicted that, in the domains of _cognition and control_ and _social connection_, participants would perceive relatively more dramatic developmental changes extending later into childhood (i.e., lower intercepts and steeper slopes), whereas in the domains of _bodily sensations_ and _negative affect_ participants would perceive relatively smaller changes across development (i.e., higher intercepts and shallower slopes).

Following Study 2, to test these hypotheses we conducted a multilevel beta regression via a generalized additive model, regressing item-level responses onto the domain and a spline-based smooth of the target age of the child fitted separately for each domain, including the maximal random effects structure that would allow the model to converge. Using the "mgcv" package for R [@wood2017], we specified this model as follows: _gamm(responses ~ s(target age in years, by = domain) + domain, random = list(subject ID = ~1, capacity = ~1)_. Again, attempting to model random slopes for domain or for target age resulted in model convergence problems. Again, responses were rescaled to range from 0.005 to 0.995, and domain was dummy-coded with _cognition and control_ as the baseline.

(We note that, in contrast to Study 2, in Study 3 we did preregister this analysis, although we preregistered a linear regression rather than a beta regression via generalized additive model; see extended notes on our choice of this analysis approach under Study 2.)

#### Preregistered exploratory analysis of perceived developmental trajectories

Beyond the two hypotheses described in the previous section, we also anticipated that perceptions of developmental trajectories in different domains would vary in their curvilinearity—e.g., development in one domain might be perceived to be gradual and linear, while development in another domain might be perceived to occur in stages, at an exponential rate, or to peak at a specific age. To explore these possibilities, we preregistered a regression analysis adding polynomial effects of target age; however, in our reconceptualization of the appropriate way to analyze our data given the distribution of the response variable (see Study 2), this exploratory analysis was superseded by the generalized additive model.

#### Preregistered exploratory analysis of developmental mechanisms

Finally, we assessed which developmental mechanisms were considered to be the most or least important in the development of each of the individual mental capacities included in this study (*bodily sensations*, *negative affect*, *social connection*, and *cognition and control*). We specified in our preregistration that this assessment would be conducted primarily through visualization, but to complement this visualization we also conducted a multilevel linear regression regressing responses to this question onto type of developmental mechanism (extrinsic vs. intrinsic), domain, and an interaction between them, including the maximal random effects structure that allowed the model to converge. Using the "lme4" package for R [@bates.etal2015], we specified this model as follows: _lmer(response ~ mechanism type * domain + (1 + mechanism type + domain | participant) + (1 | developmental mechanism) + (1 | capacity)_. We also conducted separate, analogous regression analyses for each domain separately. In all of these analyses, mechanism type and domain were effect-coded for comparisons to the grand mean collapsing across mechanism types and domains. (Note that these analyses omitted the mechanism "the child's brain changes [...]", because it was the only mechanism that was difficult to classify as either intrinsic or extrinsic via EFA and clustering analyses; see main text.)
